---
title: "Math"
author: "SA22231050"
date: "2023.12.05"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Math}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
The objective of this assignment was to familiarise myself with the use of Rcpp through the final assignment, which implemented two simple numerical methods, Newton's method for root finding and bisection to find the target value. Only the most basic functions were implemented.

## 二分法求目标值
比较目标数和数组的中间数的大小，然后将数组中的数据不断分为两份，从而判断目标数据应该处于数组的左半段还是右半段

```{r, eval=FALSE}
#R实现
bisection <- function(x,  target) {
  l = 1
  r = length(x)
  iter_n = 0
  
  ITER_UPPER = 500
  iter_n = 0
  
  while (TRUE) {
    iter_n = iter_n + 1
    mid = (l + r) / 2
    mid = floor(mid)
    if (iter_n > ITER_UPPER) {
      stop('not found solution')
    }
    mid_val = x[mid]
    if (mid_val < target) {
      l = mid+1
    } else if (mid_val > target) {
      r = mid-1
    } else {
      return (mid)
    }
  }
  return(-1)
}
x <- seq(1:100)
bisection(x, 15)

#c++
library(Rcpp)
cppFunction(code='
int binarySearchC(NumericVector x, double target) {
  int low = 0;                  
  int high = x.size() - 1;      
  
  while (low <= high) {
    int mid = (low + high) / 2; 
    double midVal = x[mid];     
    
    if (midVal < target) {
      low = mid + 1;            
    } else if (midVal > target) {
      high = mid - 1;           
    } else {
      return mid+1;               
    }
  }
')

x <- seq(1:100)
result <- binarySearchC(x, target = 15)
print(result)

#速度比较
library(microbenchmark)
tm1 <- microbenchmark::microbenchmark(
   rnR = bisection(x = seq(1:100), 15),
   rnC = binarySearchC(NumericVector x = seq(1:100), target = 15)
)
print(summary(tm1)[,c(1,3,5,6)])

```

## 牛顿法
基本牛顿法是一种是用导数的算法，它每一步的迭代方向都是沿着当前点函数值下降的方向。

$$
f\left(x\right)=f\left(x_{k}\right)+f^{\prime}\left(x_{k}\right)\left(x-x_{k}\right)+\frac12f^{\prime\prime}\left(x_{k}\right)\left(x-x_{k}\right)^{2}
$$
对上式求导并令其为0，则为
$$
f^{\prime}\left(x_{k}\right)+f^{\prime\prime}\left(x_{k}\right)(x-x_{k})=0
$$
即得到
$$x=x_{k}-\frac{f^{\prime}\left(x_{k}\right)}{f^{\prime\prime}\left(x_{k}\right)}$$

```{r, eval=FALSE}
#R
newton <- function(f, f1, x0, tol, max_iter = 1000) {
  x = x0
  f = f(x)
  f1 = f1(x)
  iter = 0
  while (TRUE) {
    if (f1 == 0) {
      stop("The derivative is 0")
    }
    x = x - f / f1
    f = f(x)
    f1 = f1(x)

    iter = iter + 1
    if ((abs(x - x0) < tol)) {
      return(c(x, iter))
    }
    x0 = x 
    if (iter > max_iter) {
      stop("cannot find solution")
    }
  }
}
ans <- newton(f = function(x) x^2 - 4, f1 = function(x) 2*x, x0 = -5, tol = 1e-6)

#c++
library(Rcpp)

cppFunction(code='
double f(double x) {
   return x*x-4;
}
')

cppFunction(code='
double f_prime(double x) {
   return 2*x;
}
')
cppFunction(code='
double newtonMethodC(double x0, double tol, int max_iter) {
  double x = x0;                 
  double fx = f(x);  
  double fpx = f_prime(x);  
  int iter = 0;               
  
  while (fabs(fx) > tol && iter < max_iter) {
    x = x - fx / fpx;         
    fx = f(x);      
    fpx = f_prime(x);  
    iter++;                    
  }
  
  return x;                    
}
')
x0 <- -9
tol <- 1e-6
max_iter <- 100
result <- newtonMethodC(x0, tol, max_iter)
print(result)


```

